---
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# False Overlapped-Cluster Rate (FOCR)

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/focr)](https://CRAN.R-project.org/package=focr)
[![R-CMD-check](https://github.com/dipterix/focr/workflows/R-CMD-check/badge.svg)](https://github.com/dipterix/focr/actions)
<!-- badges: end -->

A two-stage procedure to solve the following multiple testing problems with topological constraints:

\[
H_{0}(s): \mu(s) = 0,\quad H_{1}(s):\mu(s)\neq 0
\]

In functional data analysis, the underlying function $\mu(s)$ may be subject to topological constraints (temporal, spatial, ...). The functional domain is also uncountable. 
It is scientifically meaningful to extract blocks (clusters, or connected regions) of $s$ such that $H_{0}(s)$ are rejected. The FOCR framework controls the type-I error in the following two stages:

* Stage-I: given blocks of hypotheses, control the FOCR (an extended false cluster rate) at given level. This step is to find connected regions of interests. The FOCR is defined as follow, where $B_{k}$ are the overlapped clusters (blocks), and $D_{k}$ are the corresponding disjoint kernels;

![FOCR-Definition.svg](https://raw.githubusercontent.com/dipterix/focr/main/inst/definition.svg)

* Stage-II: given the previous rejection (clusters), calculate conditional p-values in a post-selection fashion. Then a further selection procedure will be applied to the conditional p-values to control the FDR (at individual level).

Please read [vignettes](http://github.dipterix.org/focr/articles/false-overlapped-cluster-rate.html), and [help documents](http://github.dipterix.org/focr/reference/focr.html) for more examples.

## Installation

You can install the released version of `focr` from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("focr")
```
And the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("remotes")
remotes::install_github("dipterix/focr")
```
## Example

Let's sample from this 2D image (`32x32` pixels). The underlying signal is a triangle. The noise is generated with correlation.

```{r data, message=FALSE, echo=FALSE, fig.height=4}
library(focr)
set.seed(400)
generator <- simulation_data_2D(cov_type = 'AR')
data <- generator$gen_data(0.4)
par(mfrow=c(1,2))
image(matrix(generator$mu, 32), main = 'Undelying mean')
image(matrix(colMeans(data), 32), main = 'Sample mean', sub = "Auto-correlated noise")
```

Example code:

```{r example}
res <- focr(data, block_size = 3, alpha = 0.05, 
            fdr_method = 'BH', dimension = c(32,32))
```

The initial clusters, conditional p-values, and final rejections are displayed as follows:

```{r, fig.height=3, message=FALSE, echo=FALSE}
par(mfrow=c(1,3))
mat <- matrix(0, 32, 32)
mat[res$rej_hypotheses] <- 1
image(mat, main = "FOCR initial rejections", sub = "(FOCR=0.05)")

image(1-matrix(res$cond_pvals, 32, 32), main = "Conditional p-values")

mat <- matrix(0, 32, 32)
mat[res$post_selection$rejs] <- 1
image(mat, main = "FOCR-BH final rejections", sub = "(FDR=0.05)")
```

Further reading materials:  [vignettes](http://github.dipterix.org/focr/articles/false-overlapped-cluster-rate.html), and [help documents](http://github.dipterix.org/focr/reference/focr.html).


## Citations
